Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-09-19 main.tex
minor changes

This section is devoted to geometric operations that are crucially important for our main result. 
Suppose that we have a shellable simplicial complex. 
Every time we add a simplex according to the shelling sequence, 
we complete the local star of the some simplex in the triangulation (see Lemma~\ref{lemma:existenceofstar}).
We will need a homeomorphism going from the new simplex onto its complement within the newly completed star.
This homeomorphism is required to preserve the interface and to be bi-Lipschitz.
We are particularly interested in the norms of its Jacobian.
We inerpret this as a nonlinear reflection. It will be used subsequently in generalizing the discussion in Section~\ref{section:gradient} to the setting of differential forms.

In addition to that, this endeavor will produce another geometric tool 
that enables an additional estimate Poincar\'e--Friedrichs constants:
we construct a bi-Lipschitz deformation which maps the entire star 
into the complement of the newly completed star. 
Again, we give explicit estimates of the Jacobian.
\\

@@

\begin{lemma}\label{lemma:oppositesubsimplex}
    Let $v_0, v_1, \dots, v_n \in \bbR^n$ be the vertices of a simplex $T$.
    Let $F, F'$ be the two subsimplices of $T$ with respective vertices $v_0,\dots,v_k$ and $v_{k+1},\dots,v_n$.
    Then $T$ lies on one side of the affine subspace
    \begin{gather*}
        A 
        := 
        v_0 
        + 
        \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \} 
        +  
        \operatorname{lin}\{ v_{k+1} - v_n, \dots, v_{n-1} - v_n \}.
    \end{gather*}
    Moreover, $F = T \cap A$.
\end{lemma}
\begin{proof}
    Evidently, the linear spans 
    \begin{gather*}
        A_{0} := \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \},
        \qquad 
        A_{1} := \operatorname{lin}\{ v_{k+1} - v_n, \dots, v_{n-1} - v_n \}
    \end{gather*}
    intersect only at the origin, since otherwise the vertices of $T$ are not affinely independent. 
    So $A$ has dimension $n-1$. 
    Obviously, $F \subset A \cap T$. 
    Let us now suppose that $x \in A \cap T$.
    For some real coefficients $\lambda_1,\dots,\lambda_{n-1} \in \bbR$, it holds that 
    \begin{align*}
        x &= 
        v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ) + \sum_{i=k+1}^{n-1} \lambda_{i} ( v_n - v_i )
        \\&
        = 
        \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right) v_0 
        + 
        \sum_{i=1  }^{k  } \lambda_{i} v_i 
        + 
        \left( \sum_{i=k+1}^{n-1} \lambda_{i} \right) v_n
        + 
        \sum_{i=k+1}^{n-1} (-\lambda_{i}) v_i 
        .
    \end{align*}
    We know that this is the unique expression of $x$ as non-negative linear combination of the vertices of $T$.
    Hence the $\lambda_{k+1}, \dots, \lambda_{n-1}$ are non-positive.
    But if $x \notin F$, then at least one of them must be negative,
    and thus their sum must be negative, which is a contradiction. 
    Hence the $\lambda_{k+1}, \dots, \lambda_{n-1}$ are zero,
    which implies $x \in F$.
    The desired statement follows.         
\end{proof}

@@

Next, we observe that stars around subsimplices can represented by vertex stars after a local subdivision of the mesh.
The shape regularity is controlled in that process.

@@

We know construct the main result of this section, namely the desired bi-Lipschitz homeomorphism. 
It can be imagined as the reflection across the interface between the selected simplex and the remainder of the star. 
The construction is completed in several steps,
and the Lipschitz estimate reflects the shape regularity.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    such that 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq 
        C_{\PF,T,k,p}(T_{0})
        \prod_{i=0}^{m} 
        C_{???}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    We prove this via induction. 
    The statement is clearly satisfied when $m=0$ because then we only consider a single simplex. 
    Let us assume that the statement is true for some integer $m-1 \geq 0$.

>>>>

This section is devoted to geometric operations that are crucial for our main result. 
Consider the situation where we have an $n$-dimensional local star around some simplex $S$ and some $n$-dimensional simplex $T$ within that local star. We will need a homeomorphism going from the simplex $T$ onto its complement within the local star around $S$.
This homeomorphism is required to preserve the interface and we interpret it as a nonlinear reflection. 
We ensure that the homeomorphism is bi-Lipschitz and we are particularly interested in the norms of its Jacobian.
This nonlinear reflection will be used subsequently in generalizing the discussion in Section~\ref{section:gradient} to the setting of differential forms. 
Additionally, this endeavor produces another geometric tool:
we construct a bi-Lipschitz deformation which contracts the entire star into the complement of the newly completed star.

@@

\begin{lemma}\label{lemma:oppositesubsimplex}
    Let $v_0, v_1, \dots, v_n \in \bbR^n$ be the vertices of a simplex $T$.
    Let $F, F'$ be the two subsimplices of $T$ with respective vertices $v_0,\dots,v_k$ and $v_{k+1},\dots,v_n$.
    Then $T$ lies on one side of the affine subspace
    \begin{gather*}
        A 
        := 
        v_0 
        + 
        \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \} 
        +  
        \operatorname{lin}\{ v_{k+1} - v_n, \dots, v_{n-1} - v_n \}.
    \end{gather*}
    Moreover, $F = T \cap A$.
\end{lemma}
\begin{proof}
    Evidently, the linear spans 
    \begin{gather*}
        A_{0} := \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \},
        \qquad 
        A_{1} := \operatorname{lin}\{ v_{k+1} - v_n, \dots, v_{n-1} - v_n \}
    \end{gather*}
    intersect only at the origin, since otherwise the vertices of $T$ are not affinely independent. 
    So $A$ has dimension $n-1$. 
    Obviously, $F \subset A \cap T$. 
    Suppose now that $x \in A \cap T$.
    There exist $\lambda_1,\dots,\lambda_{n-1} \in \bbR$ such that 
    \begin{align*}
        x &= 
        v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ) + \sum_{i=k+1}^{n-1} \lambda_{i} ( v_n - v_i )
        \\&
        = 
        \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right) v_0 
        + 
        \sum_{i=1  }^{k  } \lambda_{i} v_i 
        + 
        \left( \sum_{i=k+1}^{n-1} \lambda_{i} \right) v_n
        + 
        \sum_{i=k+1}^{n-1} (-\lambda_{i}) v_i 
        .
    \end{align*}
    This is the unique expression of $x$ as non-negative linear combination of the vertices of $T$.
    Hence the $\lambda_{k+1}, \dots, \lambda_{n-1}$ are non-positive.
    But if $x \notin F$, then at least one of them must be negative,
    and thus their sum must be negative, which is a contradiction. 
    Hence $\lambda_{k+1} = \dots = \lambda_{n-1} = 0$, implying $x \in F$.
\end{proof}

@@

Next, we observe local stars can be subdivided to vertex stars.
The shape regularity is controlled in that process.

@@

We now provide the bi-Lipschitz nonlinear reflection across the interface between the selected simplex and the remainder of the local star
as well as the bi-Lipschitz contraction from the local star onto the complement of the selected simplex.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior:contraction}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    such that 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq 
        C_{\PF,T,k,p}(T_{0})
        \prod_{i=0}^{m} 
        C_{???}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    We prove this via induction. 
    The statement is clearly satisfied when $m=0$ because then we only consider a single simplex. 
    Let us assume that the statement is true for some integer $m-1 \geq 0$.
